---
title: "NYC Airbnb Market Analysis"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: cosmo
    css: dashboard_styles.css
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(lubridate)
library(janitor)
library(knitr)
library(scales)
library(ggplot2)
library(plotly)
library(DT)
library(leaflet)
library(rpart)

# Charger et nettoyer les données
airbnb_raw <- read_csv("data/train.csv", show_col_types = FALSE) |>
  clean_names()

airbnb_clean <- airbnb_raw |>
  mutate(
    last_review = suppressWarnings(ymd(last_review)),
    reviews_per_month = replace_na(reviews_per_month, 0)
  ) |>
  filter(price > 0) |>
  select(-any_of(c("id", "host_id", "host_name")))

# Préparer pour la modélisation
airbnb_model <- airbnb_clean |>
  mutate(
    log_price = log1p(price),
    room_type = as.factor(room_type),
    neighbourhood_group = as.factor(neighbourhood_group),
    number_of_reviews = as.numeric(number_of_reviews)
  )

set.seed(42)
n <- nrow(airbnb_model)
idx <- sample.int(n, size = floor(0.7*n))
train <- airbnb_model[idx, ]
test  <- airbnb_model[-idx, ]

# Modèles
lm_fit <- lm(log_price ~ neighbourhood_group + room_type + number_of_reviews, data = train)
tree_fit <- rpart(
  log_price ~ neighbourhood_group + room_type + number_of_reviews,
  data = train,
  method = "anova",
  control = rpart.control(cp = 0.01, xval = 10)
)
best_cp <- tree_fit$cptable[which.min(tree_fit$cptable[, "xerror"]), "CP"]
tree_fit_pruned <- prune(tree_fit, cp = best_cp)

# Prédictions
pred_log_lm_test <- predict(lm_fit, newdata = test)
pred_price_lm_test <- expm1(pred_log_lm_test)
pred_log_tree_test <- predict(tree_fit_pruned, newdata = test)
pred_price_tree_test <- expm1(pred_log_tree_test)

# RMSE
rmse <- function(y_hat, y) sqrt(mean((y_hat - y)^2, na.rm = TRUE))
rmse_lm <- rmse(pred_price_lm_test, test$price)
rmse_tree <- rmse(pred_price_tree_test, test$price)
```

Overview {data-icon="fa-home"}
=====================================

Column {data-width=350}
-----------------------------------------------------------------------

### Key Metrics

```{r}
total_listings <- nrow(airbnb_clean)
avg_price <- mean(airbnb_clean$price)
median_price <- median(airbnb_clean$price)
total_boroughs <- n_distinct(airbnb_clean$neighbourhood_group)

valueBox(
  format(total_listings, big.mark = ","),
  caption = "Total Listings",
  icon = "fa-building",
  color = "#667eea"
)
```

### Average Price

```{r}
valueBox(
  paste0("$", round(avg_price, 0)),
  caption = "Average Price per Night",
  icon = "fa-dollar-sign",
  color = "#764ba2"
)
```

### Median Price

```{r}
valueBox(
  paste0("$", round(median_price, 0)),
  caption = "Median Price per Night",
  icon = "fa-chart-line",
  color = "#f093fb"
)
```

### Boroughs Covered

```{r}
valueBox(
  total_boroughs,
  caption = "NYC Boroughs",
  icon = "fa-map-marker-alt",
  color = "#4facfe"
)
```

Column {data-width=650}
-----------------------------------------------------------------------

### Distribution of Listings by Borough (Interactive)

```{r}
borough_data <- airbnb_clean |>
  count(neighbourhood_group) |>
  arrange(desc(n))

plot_ly(borough_data, 
        x = ~reorder(neighbourhood_group, n), 
        y = ~n,
        type = 'bar',
        marker = list(color = c('#FF385C', '#FFD23F', '#00539B', '#FF6B35', '#BC4749')),
        text = ~paste0(format(n, big.mark = ","), " listings"),
        hovertemplate = '<b>%{x}</b><br>Listings: %{y:,}<extra></extra>') |>
  layout(
    title = "",
    xaxis = list(title = "Borough"),
    yaxis = list(title = "Number of Listings"),
    plot_bgcolor = 'rgba(0,0,0,0)',
    paper_bgcolor = 'rgba(0,0,0,0)'
  )
```

### Price Distribution by Borough

```{r}
plot_ly(airbnb_clean, 
        x = ~neighbourhood_group, 
        y = ~price,
        type = 'box',
        color = ~neighbourhood_group,
        colors = c('#667eea', '#764ba2', '#f093fb', '#4facfe', '#c471f5')) |>
  layout(
    title = "",
    xaxis = list(title = "Borough"),
    yaxis = list(title = "Price ($)", type = "log"),
    showlegend = FALSE,
    plot_bgcolor = 'rgba(0,0,0,0)',
    paper_bgcolor = 'rgba(0,0,0,0)'
  )
```

Visualizations {data-icon="fa-chart-bar"}
=====================================

Column {data-width=500}
-----------------------------------------------------------------------

### Room Type Distribution

```{r}
room_data <- airbnb_clean |>
  count(room_type) |>
  mutate(pct = n / sum(n) * 100)

plot_ly(room_data, 
        labels = ~room_type, 
        values = ~n,
        type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        marker = list(colors = c('#667eea', '#764ba2', '#f093fb'),
                     line = list(color = '#FFFFFF', width = 2))) |>
  layout(
    title = "",
    showlegend = TRUE,
    plot_bgcolor = 'rgba(0,0,0,0)',
    paper_bgcolor = 'rgba(0,0,0,0)'
  )
```

### Availability vs Reviews (Interactive)

```{r}
sample_data <- airbnb_clean |>
  sample_n(min(5000, nrow(airbnb_clean)))

plot_ly(sample_data, 
        x = ~availability_365, 
        y = ~number_of_reviews,
        type = 'scatter',
        mode = 'markers',
        marker = list(size = 5, 
                     color = ~price,
                     colorscale = 'Viridis',
                     showscale = TRUE,
                     colorbar = list(title = "Price ($)")),
        text = ~paste0("Price: $", price, 
                      "<br>Available: ", availability_365, " days",
                      "<br>Reviews: ", number_of_reviews),
        hovertemplate = '%{text}<extra></extra>') |>
  layout(
    title = "",
    xaxis = list(title = "Days Available per Year"),
    yaxis = list(title = "Number of Reviews"),
    plot_bgcolor = 'rgba(0,0,0,0)',
    paper_bgcolor = 'rgba(0,0,0,0)'
  )
```

Column {data-width=500}
-----------------------------------------------------------------------

### Price Comparison by Room Type

```{r}
plot_ly(airbnb_clean, 
        x = ~room_type, 
        y = ~price,
        type = 'violin',
        color = ~room_type,
        colors = c('#667eea', '#764ba2', '#f093fb'),
        box = list(visible = TRUE),
        meanline = list(visible = TRUE)) |>
  layout(
    title = "",
    xaxis = list(title = "Room Type"),
    yaxis = list(title = "Price ($)", type = "log"),
    showlegend = FALSE,
    plot_bgcolor = 'rgba(0,0,0,0)',
    paper_bgcolor = 'rgba(0,0,0,0)'
  )
```

### Average Price by Borough and Room Type

```{r}
avg_price_data <- airbnb_clean |>
  group_by(neighbourhood_group, room_type) |>
  summarise(avg_price = mean(price), .groups = 'drop')

plot_ly(avg_price_data, 
        x = ~neighbourhood_group, 
        y = ~avg_price,
        color = ~room_type,
        colors = c('#667eea', '#764ba2', '#f093fb'),
        type = 'bar',
        text = ~paste0("$", round(avg_price, 0)),
        textposition = 'outside',
        hovertemplate = '<b>%{x}</b><br>%{data.name}<br>Avg Price: $%{y:.0f}<extra></extra>') |>
  layout(
    title = "",
    xaxis = list(title = "Borough"),
    yaxis = list(title = "Average Price ($)"),
    barmode = 'group',
    plot_bgcolor = 'rgba(0,0,0,0)',
    paper_bgcolor = 'rgba(0,0,0,0)'
  )
```

Map {data-icon="fa-map"}
=====================================

Column {data-width=1000}
-----------------------------------------------------------------------

### Interactive Map of NYC Airbnb Listings

```{r}
# Échantillon pour la carte
map_sample <- airbnb_clean |>
  sample_n(min(2000, nrow(airbnb_clean)))

# Définir les couleurs par type de chambre
color_palette <- colorFactor(
  palette = c('#667eea', '#764ba2', '#f093fb'),
  domain = map_sample$room_type
)

leaflet(map_sample) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addCircleMarkers(
    lng = ~longitude,
    lat = ~latitude,
    radius = 4,
    color = ~color_palette(room_type),
    fillOpacity = 0.6,
    stroke = TRUE,
    weight = 1,
    popup = ~paste0(
      "<strong>", name, "</strong><br>",
      "Borough: ", neighbourhood_group, "<br>",
      "Neighbourhood: ", neighbourhood, "<br>",
      "Room Type: ", room_type, "<br>",
      "Price: $", price, "<br>",
      "Reviews: ", number_of_reviews
    )
  ) |>
  addLegend(
    position = "bottomright",
    pal = color_palette,
    values = ~room_type,
    title = "Room Type",
    opacity = 0.8
  )
```

Data Table {data-icon="fa-table"}
=====================================

Column {data-width=1000}
-----------------------------------------------------------------------

### Explore All Listings (Interactive Table)

```{r}
table_data <- airbnb_clean |>
  select(
    name,
    neighbourhood_group,
    neighbourhood,
    room_type,
    price,
    minimum_nights,
    number_of_reviews,
    availability_365
  ) |>
  mutate(
    price = paste0("$", price)
  )

datatable(
  table_data,
  colnames = c(
    'Listing Name' = 'name',
    'Borough' = 'neighbourhood_group',
    'Neighbourhood' = 'neighbourhood',
    'Room Type' = 'room_type',
    'Price' = 'price',
    'Min Nights' = 'minimum_nights',
    'Reviews' = 'number_of_reviews',
    'Availability' = 'availability_365'
  ),
  filter = 'top',
  options = list(
    pageLength = 25,
    scrollX = TRUE,
    autoWidth = TRUE,
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel')
  ),
  extensions = 'Buttons',
  class = 'display compact'
)
```

Models {data-icon="fa-robot"}
=====================================

Column {data-width=400}
-----------------------------------------------------------------------

### Model Performance

```{r}
gauge(
  round(rmse_lm, 0),
  min = 0,
  max = 500,
  symbol = '$',
  label = "Linear Regression RMSE",
  gaugeSectors(
    success = c(0, 150),
    warning = c(150, 300),
    danger = c(300, 500)
  )
)
```

### Tree Model Performance

```{r}
gauge(
  round(rmse_tree, 0),
  min = 0,
  max = 500,
  symbol = '$',
  label = "Regression Tree RMSE",
  gaugeSectors(
    success = c(0, 150),
    warning = c(150, 300),
    danger = c(300, 500)
  )
)
```

### R² Score (Linear Model)

```{r}
r2_lm <- summary(lm_fit)$r.squared
gauge(
  round(r2_lm * 100, 1),
  min = 0,
  max = 100,
  symbol = '%',
  label = "R² Score",
  gaugeSectors(
    danger = c(0, 30),
    warning = c(30, 60),
    success = c(60, 100)
  )
)
```

Column {data-width=600}
-----------------------------------------------------------------------

### Model Comparison

```{r}
comparison_data <- tibble(
  Model = c("Linear Regression", "Regression Tree"),
  RMSE = c(rmse_lm, rmse_tree),
  Type = c("Linear", "Tree-based")
)

plot_ly(comparison_data, 
        x = ~Model, 
        y = ~RMSE,
        type = 'bar',
        marker = list(color = c('#667eea', '#764ba2')),
        text = ~paste0("$", round(RMSE, 0)),
        textposition = 'outside',
        hovertemplate = '<b>%{x}</b><br>RMSE: $%{y:.2f}<extra></extra>') |>
  layout(
    title = "RMSE Comparison (Lower is Better)",
    xaxis = list(title = ""),
    yaxis = list(title = "RMSE ($)"),
    plot_bgcolor = 'rgba(0,0,0,0)',
    paper_bgcolor = 'rgba(0,0,0,0)'
  )
```

### Variable Importance (Regression Tree)

```{r}
importance_data <- tibble(
  Variable = names(tree_fit_pruned$variable.importance),
  Importance = tree_fit_pruned$variable.importance
) |>
  mutate(Share = Importance / sum(Importance) * 100) |>
  arrange(desc(Share))

plot_ly(importance_data, 
        x = ~Share, 
        y = ~reorder(Variable, Share),
        type = 'bar',
        orientation = 'h',
        marker = list(color = '#764ba2'),
        text = ~paste0(round(Share, 1), "%"),
        textposition = 'outside',
        hovertemplate = '<b>%{y}</b><br>Importance: %{x:.1f}%<extra></extra>') |>
  layout(
    title = "",
    xaxis = list(title = "Relative Importance (%)"),
    yaxis = list(title = ""),
    plot_bgcolor = 'rgba(0,0,0,0)',
    paper_bgcolor = 'rgba(0,0,0,0)'
  )
```

Column {data-width=400}
-----------------------------------------------------------------------

### Coefficient Effects (Linear Regression)

```{r}
coefs <- coefficients(lm_fit)
effects_price <- expm1(coefs)
percentOfInfluence <- c(
  effects_price["neighbourhood_groupBrooklyn"] * 100,
  effects_price["neighbourhood_groupManhattan"] * 100,
  effects_price["neighbourhood_groupQueens"] * 100,
  effects_price["neighbourhood_groupStaten Island"] * 100,
  effects_price["room_typePrivate room"] * 100,
  effects_price["room_typeShared room"] * 100
)

coef_data <- tibble(
  Variable = c(
    "Brooklyn",
    "Manhattan",
    "Queens",
    "Staten Island",
    "Private Room",
    "Shared Room"
  ),
  Effect = percentOfInfluence
) |>
  arrange(desc(Effect))

plot_ly(coef_data, 
        x = ~Effect, 
        y = ~reorder(Variable, Effect),
        type = 'bar',
        orientation = 'h',
        marker = list(color = ~ifelse(Effect > 0, '#667eea', '#f093fb')),
        text = ~paste0(ifelse(Effect > 0, "+", ""), round(Effect, 1), "%"),
        textposition = 'outside',
        hovertemplate = '<b>%{y}</b><br>Effect: %{x:.2f}%<extra></extra>') |>
  layout(
    title = "Price Impact by Factor",
    xaxis = list(title = "Effect on Price (%)"),
    yaxis = list(title = ""),
    plot_bgcolor = 'rgba(0,0,0,0)',
    paper_bgcolor = 'rgba(0,0,0,0)'
  )
```

About {data-icon="fa-info-circle" style="position:relative;"}
=====================================

Column {data-width=1000 style="overflow-y:auto;"} 
-----------------------------------------------------------------------

### About This Dashboard

<div style="padding: 30px; background: linear-gradient(135deg, #667eea 0%, #764ba2 100%); border-radius: 10px; color: white; margin: 20px;">

<h2 style="color: white; font-weight: bold;">NYC Airbnb Market Analysis Dashboard</h2>

<p style="font-size: 1.1em; line-height: 1.8;">
This interactive dashboard explores the dynamics of the Airbnb market in New York City. It provides insights into:
</p>

<ul style="font-size: 1.05em; line-height: 2;">
  <li><strong>Market Overview:</strong> Key metrics and distribution patterns across NYC boroughs</li>
  <li><strong>Visualizations:</strong> Interactive charts showing price patterns, room types, and availability</li>
  <li><strong>Geographic Analysis:</strong> Interactive map displaying listings across the city</li>
  <li><strong>Data Exploration:</strong> Searchable and filterable table of all listings</li>
  <li><strong>Predictive Models:</strong> Machine learning models comparing Linear Regression and Regression Trees</li>
</ul>

<h3 style="color: white; margin-top: 30px;">Dataset Information</h3>
<p style="font-size: 1.05em;">
The dataset contains <strong>`r format(nrow(airbnb_clean), big.mark = ",")`</strong> Airbnb listings across NYC's five boroughs, with information on pricing, location, room type, and host activity.
</p>

<h3 style="color: white; margin-top: 30px;">Authors</h3>
<p style="font-size: 1.1em; font-weight: 600;">
Antoine, Charles, Mokhmad
</p>

<p style="font-size: 0.9em; margin-top: 30px; opacity: 0.8;">
</p>

</div>

### Key Findings {.scrollable-content}

<div class="info-card scrollable-section" style="padding: 25px; background: white; border-radius: 10px; border-left: 5px solid #667eea; margin: 20px; box-shadow: 0 4px 15px rgba(0,0,0,0.1);">

<h4 style="color: #667eea;">📊 Market Insights</h4>

<ul style="line-height: 2;">
  <li><strong>Manhattan and Brooklyn</strong> dominate the market with the highest concentration of listings</li>
  <li><strong>Entire homes/apartments</strong> represent the most common room type</li>
  <li><strong>Median price: $`r round(median_price, 0)`</strong> per night across all boroughs</li>
  <li><strong>Price variation</strong> is significant within boroughs, with some neighborhoods commanding premium rates</li>
</ul>

<h4 style="color: #764ba2; margin-top: 25px;">🤖 Model Performance</h4>

<ul style="line-height: 2;">
  <li>Both models achieve similar RMSE of approximately <strong>$`r round(rmse_lm, 0)`</strong></li>
  <li><strong>Room type</strong> is the most important predictor (~75% importance in tree model)</li>
  <li><strong>Location</strong> (borough/neighborhood) accounts for ~22% of price variation</li>
  <li>The models explain <strong>`r round(summary(lm_fit)$r.squared * 100, 1)`%</strong> of price variance</li>
</ul>

</div>