---
title: "NYC Airbnb Market Analysis"
author: "Antoine, Charles, Mokhmad"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    theme: readable
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: true
    toc_depth: 1
    css: styles.css
    df_print: paged
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6,
  fig.align = 'center'
)

# Libraries
library(tidyverse)
library(lubridate)
library(janitor)
library(knitr)
library(scales)
library(ggplot2)
library(leaflet)
library(shiny)
library(bslib)
library(rpart)
```

<div id="introduction" class="info-box">

## Introduction {-}

This report explores the dynamics of the **Airbnb market in New York City** using a public dataset. The objective is to better understand price patterns, geographical trends, and the factors that influence rental values across the city.

**Key Questions:**

- How are listings distributed across NYC boroughs?
- What factors drive price variations?
- Can we predict Airbnb prices with machine learning?

</div>

---

# Understanding the Dataset {.tabset .tabset-fade .tabset-pills}

## Dataset Overview

The dataset gathers detailed information about Airbnb listings across New York City. It covers variables related to pricing, location, property type, and host activity, providing a rich foundation to understand market dynamics.

Before any analysis, the dataset is inspected to assess its structure and scope.

```{r}
# Import and clean column names
airbnb_raw = read_csv("data/train.csv", show_col_types = FALSE) |>
  clean_names()

# Dimensions of the dataset
n_rows = nrow(airbnb_raw)
n_cols = ncol(airbnb_raw)

kable(
  tibble(
    "Number of observations" = format(n_rows, big.mark = ","),
    "Number of variables" = n_cols
  ),
  caption = "Dataset overview: size and scope",
  align = 'c'
)
```

## Variables

```{r}
# Display column names
kable(
  tibble("Column names" = colnames(airbnb_raw)),
  caption = "Variables available in the dataset"
)
```

## Data Cleaning

<div class="highlight-box">

To ensure a reliable foundation for the analysis, several adjustments were made:

- **Temporal data**: the variable `last_review` was converted into a valid date format to allow time-based exploration.
- **Missing information**: listings with no reviews had missing values for `reviews_per_month`; these were replaced by 0, meaning no activity.
- **Unrealistic values**: listings with a price of 0 were removed, as such prices are not representative of the market.
- **Non-relevant identifiers**: variables like `id`, `host_id`, and `host_name` were excluded to focus on analytical information.

</div>

After these steps, the dataset is now consistent, cleaned, and ready for exploration.

```{r}
# Apply all cleaning steps
airbnb_clean = airbnb_raw |>
  mutate(
    last_review = suppressWarnings(ymd(last_review)),
    reviews_per_month = replace_na(reviews_per_month, 0)
  ) |>
  filter(price > 0) |>
  select(-any_of(c("id", "host_id", "host_name")))

# Display final dimensions after cleaning
kable(
  tibble(
    "Observations after cleaning" = format(nrow(airbnb_clean), big.mark = ","),
    "Variables after cleaning" = ncol(airbnb_clean)
  ),
  caption = "Dataset after cleaning and preparation",
  align = 'c'
)
```

# {-}

---

# Market Overview {.tabset .tabset-fade .tabset-pills}

## General Distribution

<div class="result-card">

The cleaned dataset now provides a clear view of the Airbnb market across New York City. This section explores the overall composition of the listings and the main factors that shape price dynamics.

</div>

Most listings are concentrated in **Manhattan** and **Brooklyn**, which also represent the most active and expensive areas in the city. Understanding how listings are distributed geographically helps identify the core zones of Airbnb activity.

```{r}
# Distribution by borough
borough_summary = airbnb_clean |>
  group_by(neighbourhood_group) |>
  summarise(
    listings = n(),
    avg_price = mean(price, na.rm = TRUE)
  ) |>
  arrange(desc(listings)) |>
  mutate(
    listings = format(listings, big.mark = ","),
    avg_price = paste0("$", round(avg_price, 2))
  )

kable(
  borough_summary, 
  caption = "Distribution of listings and average price by borough",
  col.names = c("Borough", "Number of Listings", "Average Price"),
  align = 'lcc'
)
```

## Availability & Demand

Listings with **fewer available days** usually have more reviews, showing higher demand and frequent bookings. Those available all year but with few reviews seem less attractive to guests.

```{r, fig.height=5}
ggplot(airbnb_clean, aes(x = availability_365, y = number_of_reviews)) +
  geom_hex(bins = 30) +
  scale_fill_viridis_c(option = "plasma") +
  labs(
    title = "Relationship between Availability and Number of Reviews",
    x = "Days Available per Year",
    y = "Number of Reviews",
    fill = "Density"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    panel.grid.minor = element_blank()
  )
```

## Room Types

<div class="success-box">

**Entire homes** and **private rooms** dominate the market, showing that most hosts rent full apartments or personal spaces rather than shared ones.

</div>

```{r, fig.height=4}
room_counts <- airbnb_clean |>
  count(room_type) |>
  mutate(pct = n / sum(n) * 100)

ggplot(room_counts, aes(x = reorder(room_type, n), y = n, fill = room_type)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = paste0(format(n, big.mark = ","), "\n(", round(pct, 1), "%)")), 
            hjust = -0.1, size = 4, fontface = "bold") +
  coord_flip() +
  scale_fill_manual(values = c("#FF385C", "#FFD23F", "#00539B")) +
  labs(
    title = "Distribution of Listings by Room Type",
    x = NULL,
    y = "Number of Listings"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 11)
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

## Price by Area

Manhattan and Brooklyn show the **highest median prices**, confirming their position as the most expensive and in-demand areas of the city.

```{r, fig.height=5}
ggplot(airbnb_clean, aes(x = reorder(neighbourhood_group, price, FUN = median), 
                          y = price, fill = neighbourhood_group)) +
  geom_boxplot(outlier.alpha = 0.3, outlier.size = 1) +
  scale_y_log10(labels = dollar_format()) +
  scale_fill_manual(values = c("#FF385C", "#FFD23F", "#00539B", "#FF6B35", "#BC4749")) +
  labs(
    title = "Distribution of Prices by Borough (Log Scale)",
    x = "Borough",
    y = "Price (log scale)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 11)
  )
```

# {-}

---

# Neighbourhood Analysis {.tabset .tabset-fade .tabset-pills}

These boxplots show how prices vary inside each borough. Even within the same area, some neighbourhoods stand out with much higher prices, especially in Manhattan and parts of Brooklyn.

## Bronx

```{r, fig.height=6}
df <- subset(airbnb_clean, neighbourhood_group == "Bronx")
if (nrow(df) > 0) {
  ggplot(df, aes(x = reorder(neighbourhood, price, FUN = median), y = price)) +
    geom_boxplot(fill = "#FF385C", outlier.alpha = 0.2, outlier.size = 0.8) +
    scale_y_log10(labels = dollar_format()) +
    coord_flip() +
    labs(
      title = "Price distribution — Bronx (log scale)",
      x = NULL,
      y = "Price (log)"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(face = "bold", size = 12),
      axis.text.y = element_text(size = 8),
      panel.grid.major.y = element_blank()
    )
}
```

## Brooklyn

```{r, fig.height=6}
df <- subset(airbnb_clean, neighbourhood_group == "Brooklyn")
if (nrow(df) > 0) {
  ggplot(df, aes(x = reorder(neighbourhood, price, FUN = median), y = price)) +
    geom_boxplot(fill = "#FFD23F", outlier.alpha = 0.2, outlier.size = 0.8) +
    scale_y_log10(labels = dollar_format()) +
    coord_flip() +
    labs(
      title = "Price distribution — Brooklyn (log scale)",
      x = NULL,
      y = "Price (log)"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(face = "bold", size = 12),
      axis.text.y = element_text(size = 8),
      panel.grid.major.y = element_blank()
    )
}
```

## Manhattan

```{r, fig.height=6}
df <- subset(airbnb_clean, neighbourhood_group == "Manhattan")
if (nrow(df) > 0) {
  ggplot(df, aes(x = reorder(neighbourhood, price, FUN = median), y = price)) +
    geom_boxplot(fill = "#00539B", outlier.alpha = 0.2, outlier.size = 0.8) +
    scale_y_log10(labels = dollar_format()) +
    coord_flip() +
    labs(
      title = "Price distribution — Manhattan (log scale)",
      x = NULL,
      y = "Price (log)"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(face = "bold", size = 12),
      axis.text.y = element_text(size = 8),
      panel.grid.major.y = element_blank()
    )
}
```

## Queens

```{r, fig.height=6}
df <- subset(airbnb_clean, neighbourhood_group == "Queens")
if (nrow(df) > 0) {
  ggplot(df, aes(x = reorder(neighbourhood, price, FUN = median), y = price)) +
    geom_boxplot(fill = "#FF6B35", outlier.alpha = 0.2, outlier.size = 0.8) +
    scale_y_log10(labels = dollar_format()) +
    coord_flip() +
    labs(
      title = "Price distribution — Queens (log scale)",
      x = NULL,
      y = "Price (log)"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(face = "bold", size = 12),
      axis.text.y = element_text(size = 8),
      panel.grid.major.y = element_blank()
    )
}
```

## Staten Island

```{r, fig.height=6}
df <- subset(airbnb_clean, neighbourhood_group == "Staten Island")
if (nrow(df) > 0) {
  ggplot(df, aes(x = reorder(neighbourhood, price, FUN = median), y = price)) +
    geom_boxplot(fill = "#BC4749", outlier.alpha = 0.2, outlier.size = 0.8) +
    scale_y_log10(labels = dollar_format()) +
    coord_flip() +
    labs(
      title = "Price distribution — Staten Island (log scale)",
      x = NULL,
      y = "Price (log)"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(face = "bold", size = 12),
      axis.text.y = element_text(size = 8),
      panel.grid.major.y = element_blank()
    )
}
```

# {-}

---

# Predictive Modeling {.tabset .tabset-fade .tabset-pills}

<div class="result-card" style="background: linear-gradient(135deg, #fff5f7 0%, #ffe0e6 100%); border-top: 4px solid #FF385C;">

This final part focuses on **predicting Airbnb prices** based on location, room type, and customer activity.

Two models are tested:

- **Linear Regression** 📊
- **Regression Tree** 🌳

We compare their accuracy and understand which factors most influence price variation across New York City.

</div>

## Data Preparation

The dataset is now prepared for modeling, with key variables formatted and the data split into **training (70%)** and **test sets (30%)** to build and evaluate price prediction models.

```{r}
airbnb_clean <- airbnb_clean |>
  mutate(
    log_price = log1p(price),
    room_type = as.factor(room_type),
    neighbourhood_group = as.factor(neighbourhood_group),
    number_of_reviews = as.numeric(number_of_reviews)
  )

set.seed(42)
n <- nrow(airbnb_clean)
idx <- sample.int(n, size = floor(0.7*n))
train <- airbnb_clean[idx, ]
test  <- airbnb_clean[-idx, ]
```

<div class="metric">
  <span class="metric-value">`r format(nrow(train), big.mark = ",")`</span>
  <span class="metric-label">Training Set</span>
</div>

<div class="metric">
  <span class="metric-value">`r format(nrow(test), big.mark = ",")`</span>
  <span class="metric-label">Test Set</span>
</div>

## Linear Regression

A **linear regression model** is trained to estimate how each factor — location, room type, and number of reviews — impacts the price.

Positive coefficients indicate a price increase, while negative ones show a decrease relative to the baseline.

```{r}
lm_fit <- lm(log_price ~ neighbourhood_group + room_type + number_of_reviews, data = train)
lm_sum <- summary(lm_fit)
```

<div class="success-box">

**Model R²:** `r round(lm_sum$r.squared, 4)` — This means the model explains **`r round(lm_sum$r.squared * 100, 2)`%** of the price variation.

</div>

### Coefficients Analysis

To better interpret the linear regression model, we translate each coefficient into a **percentage impact on price**. Positive values indicate factors that increase the average price, while negative ones represent elements that tend to lower it.

```{r}
coefs <- coefficients(lm_fit)
effects_price <- expm1(coefs)
percentOfInfluence <- c(
  effects_price["neighbourhood_groupBrooklyn"] * 100,
  effects_price["neighbourhood_groupManhattan"] * 100,
  effects_price["neighbourhood_groupQueens"] * 100,
  effects_price["neighbourhood_groupStaten Island"] * 100,
  effects_price["room_typePrivate room"] * 100,
  effects_price["room_typeShared room"] * 100
)

results <- tibble::tibble(
  Variable = c(
    "Brooklyn",
    "Manhattan",
    "Queens",
    "Staten Island",
    "Private Room",
    "Shared Room"
  ),
  `Influence on Price (%)` = round(percentOfInfluence, 2)
)

knitr::kable(
  results, 
  caption = "Estimated percentage influence of each factor on Airbnb price",
  align = 'lc'
)
```

## Regression Tree

We train a **regression tree** to capture non-linear price patterns and interactions between location and room type. The tree is pruned using cross-validation to prevent overfitting and keep only the splits that improve out-of-sample accuracy.

```{r}
# Train and prune regression tree
tree_fit <- rpart(
  log_price ~ neighbourhood_group + room_type + number_of_reviews,
  data = train,
  method = "anova",
  control = rpart.control(cp = 0.01, xval = 10)
)

best_cp <- tree_fit$cptable[which.min(tree_fit$cptable[, "xerror"]), "CP"]
tree_fit_pruned <- prune(tree_fit, cp = best_cp)
```

### Variable Importance

The regression tree provides a more **visual and hierarchical way** to understand which variables most influence price predictions.

The variable importance values highlight which factors — such as location or room type — play the biggest role in determining Airbnb prices.

```{r}
importance_tbl <- tibble::tibble(
  Variable = names(tree_fit_pruned$variable.importance),
  Importance = tree_fit_pruned$variable.importance
) |>
  mutate(Share = round(Importance / sum(Importance) * 100, 1)) |>
  arrange(desc(Share))

knitr::kable(
  importance_tbl |> select(Variable, `Relative Importance (%)` = Share),
  caption = "Relative influence of each factor in the regression tree model",
  align = 'lc'
)
```

<div class="success-box">

**Key Finding:** The regression tree shows that **room type explains about 75%** of the model's predictive power, followed by neighbourhood (~22%) and number of reviews (~3%).

This means that Airbnb prices in New York are mainly driven by the **type of accommodation** and its **location**, while guest activity (reviews) has a marginal impact.

</div>

## Model Evaluation

Both models are now tested on unseen data to evaluate how well they can predict Airbnb prices based on key factors such as location, room type, and customer activity.

```{r}
# LM
pred_log_lm_test  <- predict(lm_fit, newdata = test)
pred_price_lm_test <- expm1(pred_log_lm_test)

# Tree (pruned)
pred_log_tree_test  <- predict(tree_fit_pruned, newdata = test)
pred_price_tree_test <- expm1(pred_log_tree_test)
```

### Root Mean Squared Error (RMSE)

To evaluate model accuracy, we use the **Root Mean Squared Error (RMSE)**, which measures how far predicted prices deviate from real Airbnb prices.

**Lower values indicate better performance** and more reliable predictions.

```{r}
rmse <- function(y_hat, y) sqrt(mean((y_hat - y)^2, na.rm = TRUE))

rmse_lm   <- rmse(pred_price_lm_test,   test$price)
rmse_tree <- rmse(pred_price_tree_test, test$price)

results <- tibble::tibble(
  Model = c("Linear Regression", "Regression Tree"),
  RMSE = round(c(rmse_lm, rmse_tree), 2)
)

knitr::kable(
  results, 
  caption = "Model performance comparison (lower RMSE = better)",
  align = 'lc'
)
```

<div class="result-card">

<div class="metric">
  <span class="metric-value">$`r round(rmse_lm, 0)`</span>
  <span class="metric-label">Linear Regression RMSE</span>
</div>

<div class="metric">
  <span class="metric-value">$`r round(rmse_tree, 0)`</span>
  <span class="metric-label">Regression Tree RMSE</span>
</div>

</div>

<div class="highlight-box">

**Interpretation:** The two regression models both show an RMSE of around **$260**, indicating a rather large prediction error. With a first quartile of $69 and a median of $106, the gap between predicted and actual prices remains significant.

This suggests that both models capture only part of the market complexity and would need improvement to achieve acceptable accuracy levels.

</div>

# {-}

---

<div style="text-align: center; padding: 40px 0; color: #FF385C; font-size: 0.9em;">

**NYC Airbnb Market Analysis** | `r format(Sys.Date(), '%Y')` | Antoine, Charles, Mokhmad

</div>