---
title: "New York City Airbnb Market Analysis"
author: "Authors: Antoine, Charles, Mokhmad"
subtitle: "Strategic Insights & Predictive Modeling"
output: 
  html_document:
    theme: default
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: true
    toc_depth: 1
    number_sections: true
    css: styles.css
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 12,
  fig.height = 7,
  fig.align = 'center',
  dpi = 300
)

# Load required libraries
library(tidyverse)
library(lubridate)
library(janitor)
library(knitr)
library(scales)
library(ggplot2)
library(leaflet)
library(rpart)
library(plotly)

# Set ggplot theme for professional charts
theme_set(theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0, color = "#1a2332"),
    plot.author = element_text(face = "bold", size = 7, hjust = 0, color = "#d4d4d4"),
    plot.subtitle = element_text(size = 11, color = "#525252", margin = margin(b = 15)),
    axis.title = element_text(face = "bold", size = 10),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "#e5e5e5", size = 0.3)
  ))
```

<div class="info-box">

## Executive Summary {-}

This report presents a comprehensive analysis of the **New York City Airbnb market**, examining pricing dynamics, geographical distribution patterns, and demand indicators across all five boroughs. The study leverages a dataset of `r format(nrow(read_csv("data/train.csv", show_col_types = FALSE)), big.mark = ",")` active listings to deliver actionable insights for:

**Strategic Objectives:**

- **Market Segmentation:** Identify high-value areas and property types
- **Pricing Intelligence:** Understand factors driving price variations
- **Predictive Analytics:** Deploy machine learning models for price forecasting
- **Investment Insights:** Guide strategic decisions for property acquisitions

**Key Deliverables:** Market overview, competitive landscape analysis, and predictive modeling framework with performance benchmarks.

</div>

---

# Data Foundation & Methodology

## Dataset Characteristics

The analysis utilizes a comprehensive dataset capturing the NYC Airbnb ecosystem. This section establishes the data foundation and ensures analytical rigor through systematic quality control.

```{r}
# Import and standardize data
airbnb_raw <- read_csv("data/train.csv", show_col_types = FALSE) |>
  clean_names()

# Calculate key metrics
n_rows <- nrow(airbnb_raw)
n_cols <- ncol(airbnb_raw)

# Present dataset overview
kable(
  tibble(
    Metric = c("Total Observations", "Variables Tracked", "Data Completeness"),
    Value = c(
      format(n_rows, big.mark = ","),
      n_cols,
      "High"
    )
  ),
  caption = "**Table 1.1** — Dataset Overview & Scope",
  align = 'lr'
)
```

### Variable Inventory

```{r}
# Create professional variable table
var_summary <- tibble(
  Variable = colnames(airbnb_raw),
  Category = case_when(
    Variable %in% c("price", "minimum_nights") ~ "Financial",
    Variable %in% c("latitude", "longitude", "neighbourhood_group", "neighbourhood") ~ "Geographic",
    Variable %in% c("number_of_reviews", "reviews_per_month", "last_review") ~ "Performance",
    Variable %in% c("room_type", "availability_365") ~ "Property",
    TRUE ~ "Identifier"
  )
) |>
  arrange(Category, Variable)

kable(
  var_summary,
  caption = "**Table 1.2** — Variable Classification & Data Structure",
  align = 'll'
)
```

## Data Quality Assurance

<div class="highlight-box">

**Quality Control Protocol:**

The following data preparation procedures were implemented to ensure analytical integrity:

1. **Temporal Standardization:** Converted `last_review` to ISO 8601 date format for time-series analysis
2. **Missing Value Treatment:** Imputed zero values for `reviews_per_month` where no activity was recorded
3. **Outlier Management:** Removed listings with zero pricing (n = `r sum(read_csv("data/train.csv", show_col_types = FALSE)$price == 0, na.rm = TRUE)`) as non-representative
4. **Variable Optimization:** Excluded non-analytical identifiers (`id`, `host_id`, `host_name`)

</div>

```{r}
# Execute data cleaning pipeline
airbnb_clean <- airbnb_raw |>
  mutate(
    last_review = suppressWarnings(ymd(last_review)),
    reviews_per_month = replace_na(reviews_per_month, 0)
  ) |>
  filter(price > 0) |>
  select(-any_of(c("id", "host_id", "host_name")))

# Calculate retention rate
retention_rate <- round((nrow(airbnb_clean) / n_rows) * 100, 1)

# Present cleaned dataset metrics
kable(
  tibble(
    Metric = c("Cleaned Observations", "Variables Retained", "Data Retention Rate"),
    Value = c(
      format(nrow(airbnb_clean), big.mark = ","),
      ncol(airbnb_clean),
      paste0(retention_rate, "%")
    )
  ),
  caption = "**Table 1.3** — Post-Processing Dataset Characteristics",
  align = 'lr'
)
```

---

# Market Landscape Analysis

## Borough-Level Performance Metrics

The New York City market demonstrates significant geographical heterogeneity. Manhattan and Brooklyn emerge as dominant markets, accounting for the majority of active listings and commanding premium pricing.

```{r}
# Calculate comprehensive borough statistics
borough_metrics <- airbnb_clean |>
  group_by(neighbourhood_group) |>
  summarise(
    `Active Listings` = n(),
    `Mean Price` = mean(price, na.rm = TRUE),
    `Median Price` = median(price, na.rm = TRUE),
    `Price Std Dev` = sd(price, na.rm = TRUE),
    `Avg Reviews` = mean(number_of_reviews, na.rm = TRUE)
  ) |>
  arrange(desc(`Active Listings`)) |>
  mutate(
    `Market Share` = paste0(round(`Active Listings` / sum(`Active Listings`) * 100, 1), "%"),
    `Active Listings` = format(`Active Listings`, big.mark = ","),
    `Mean Price` = paste0("$", round(`Mean Price`, 0)),
    `Median Price` = paste0("$", round(`Median Price`, 0)),
    `Price Std Dev` = paste0("$", round(`Price Std Dev`, 0)),
    `Avg Reviews` = round(`Avg Reviews`, 1)
  )

kable(
  borough_metrics,
  caption = "**Table 2.1** — Borough Performance Dashboard: Listings, Pricing & Engagement",
  col.names = c("Borough", "Listings", "Mean", "Median", "Std Dev", "Avg Reviews", "Market Share"),
  align = 'lcccccc'
)
```

### Market Concentration Analysis

```{r, fig.height=6}
# Calculate market share data
market_share <- airbnb_clean |>
  count(neighbourhood_group) |>
  mutate(
    pct = n / sum(n) * 100,
    label = paste0(neighbourhood_group, "\n", format(n, big.mark = ","), " listings\n(", round(pct, 1), "%)")
  ) |>
  arrange(desc(n))

# Create professional bar chart
ggplot(market_share, aes(x = reorder(neighbourhood_group, n), y = n, fill = neighbourhood_group)) +
  geom_col(width = 0.7, alpha = 0.9) +
  geom_text(aes(label = format(n, big.mark = ",")), 
            hjust = -0.1, size = 4, fontface = "bold", color = "#1a2332") +
  coord_flip() +
  scale_fill_manual(values = c("#FF385C", "#FFD23F", "#00539B", "#FF6B35", "#BC4749")) +
  scale_y_continuous(labels = label_comma(), expand = expansion(mult = c(0, 0.12))) +
  labs(
    title = "Market Distribution by Borough",
    subtitle = "Active Airbnb listings across New York City boroughs",
    x = NULL,
    y = "Number of Active Listings",
    caption = "Figure 2.1"
  ) +
  theme(legend.position = "none")
```

## Property Type Segmentation

<div class="success-box">

**Market Composition Finding:**

The market is predominantly composed of **entire homes/apartments** and **private rooms**, which together represent over 98% of all listings. Shared rooms constitute a negligible market segment, indicating strong consumer preference for privacy and exclusive space access.

</div>

```{r, fig.height=5}
# Room type analysis with business metrics
room_analysis <- airbnb_clean |>
  group_by(room_type) |>
  summarise(
    count = n(),
    avg_price = mean(price, na.rm = TRUE),
    avg_reviews = mean(number_of_reviews, na.rm = TRUE)
  ) |>
  mutate(
    pct = count / sum(count) * 100
  ) |>
  arrange(desc(count))

# Visualization
ggplot(room_analysis, aes(x = reorder(room_type, count), y = count, fill = room_type)) +
  geom_col(width = 0.6, alpha = 0.9) +
  geom_text(
    aes(label = paste0(format(count, big.mark = ","), "\n", round(pct, 1), "%")),
    hjust = -0.1, size = 4, fontface = "bold", color = "#1a2332"
  ) +
  coord_flip() +
  scale_fill_manual(values = c("#FF385C", "#FFD23F", "#00539B")) +
  scale_y_continuous(labels = label_comma(), expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Property Type Distribution",
    subtitle = "Market segmentation by accommodation type",
    x = NULL,
    y = "Number of Listings",
    caption = "Figure 2.2"
  ) +
  theme(legend.position = "none")
```

## Pricing Intelligence

### Cross-Borough Price Benchmarking

```{r, fig.height=6}
# Professional boxplot for price distribution
ggplot(airbnb_clean, aes(x = reorder(neighbourhood_group, price, FUN = median), 
                          y = price, fill = neighbourhood_group)) +
  geom_boxplot(outlier.alpha = 0.3, outlier.size = 0.8, alpha = 0.85) +
  scale_y_log10(
    labels = dollar_format(),
    breaks = c(50, 100, 200, 500, 1000, 2000, 5000)
  ) +
  scale_fill_manual(values = c("#FF385C", "#FFD23F", "#00539B", "#FF6B35", "#BC4749")) +
  labs(
    title = "Price Distribution by Borough",
    subtitle = "Logarithmic scale reveals pricing tiers across NYC neighborhoods (median indicated by center line)",
    x = NULL,
    y = "Nightly Rate (USD, log scale)",
    caption = "Figure 2.3"
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 0, hjust = 0.5, face = "bold")
  ) +
  coord_flip()
```

### Demand-Availability Dynamics

```{r, fig.height=6}
# Sample data for performance visualization
sample_data <- airbnb_clean |>
  sample_n(min(5000, nrow(airbnb_clean)))

# Hexbin density plot
ggplot(sample_data, aes(x = availability_365, y = number_of_reviews)) +
  geom_hex(bins = 35, alpha = 0.9) +
  scale_fill_viridis_c(
    option = "plasma",
    name = "Listing\nDensity",
    labels = label_comma()
  ) +
  labs(
    title = "Availability-Performance Correlation",
    subtitle = "Relationship between annual availability and review volume (5,000 listing sample)",
    x = "Days Available per Year",
    y = "Total Reviews",
    caption = "Figure 2.4"
  ) +
  theme(
    legend.position = "right",
    legend.key.height = unit(1, "cm")
  )
```

<div class="highlight-box">

**Strategic Insight:**

Listings with **lower availability** (< 180 days/year) demonstrate significantly higher review counts, indicating superior occupancy rates and market demand. This inverse correlation suggests that properties with limited availability operate at premium positioning with consistent booking patterns.

</div>

---

# Neighborhood Competitive Intelligence

This section provides granular analysis at the neighborhood level, revealing intra-borough pricing variations that inform micro-market investment strategies.

## Manhattan — Premium Market Analysis

```{r, fig.height=8}
manhattan_data <- airbnb_clean |>
  filter(neighbourhood_group == "Manhattan") |>
  group_by(neighbourhood) |>
  filter(n() >= 10) |>
  ungroup()

if (nrow(manhattan_data) > 0) {
  ggplot(manhattan_data, aes(x = reorder(neighbourhood, price, FUN = median), y = price)) +
    geom_boxplot(fill = "#00539B", alpha = 0.8, outlier.alpha = 0.2) +
    scale_y_log10(labels = dollar_format()) +
    coord_flip() +
    labs(
      title = "Manhattan Neighborhood Price Benchmarking",
      subtitle = "Median nightly rates by neighborhood (minimum 10 listings)",
      x = NULL,
      y = "Nightly Rate (USD, log scale)",
      caption = "Figure 3.1"
    ) +
    theme(axis.text.y = element_text(size = 7))
}
```

## Brooklyn — Emerging Market Opportunities

```{r, fig.height=8}
brooklyn_data <- airbnb_clean |>
  filter(neighbourhood_group == "Brooklyn") |>
  group_by(neighbourhood) |>
  filter(n() >= 10) |>
  ungroup()

if (nrow(brooklyn_data) > 0) {
  ggplot(brooklyn_data, aes(x = reorder(neighbourhood, price, FUN = median), y = price)) +
    geom_boxplot(fill = "#FFD23F", alpha = 0.8, outlier.alpha = 0.2) +
    scale_y_log10(labels = dollar_format()) +
    coord_flip() +
    labs(
      title = "Brooklyn Neighborhood Value Analysis",
      subtitle = "Price distribution reveals growth corridors and value segments",
      x = NULL,
      y = "Nightly Rate (USD, log scale)",
      caption = "Figure 3.2"
    ) +
    theme(axis.text.y = element_text(size = 7))
}
```

## Queens — Diverse Market Landscape

```{r, fig.height=8}
queens_data <- airbnb_clean |>
  filter(neighbourhood_group == "Queens") |>
  group_by(neighbourhood) |>
  filter(n() >= 10) |>
  ungroup()

if (nrow(queens_data) > 0) {
  ggplot(queens_data, aes(x = reorder(neighbourhood, price, FUN = median), y = price)) +
    geom_boxplot(fill = "#FF6B35", alpha = 0.8, outlier.alpha = 0.2) +
    scale_y_log10(labels = dollar_format()) +
    coord_flip() +
    labs(
      title = "Queens Neighborhood Market Analysis",
      subtitle = "Competitive pricing across diverse neighborhoods",
      x = NULL,
      y = "Nightly Rate (USD, log scale)",
      caption = "Figure 3.3"
    ) +
    theme(axis.text.y = element_text(size = 7))
}
```

## Bronx — Value Market Segment

```{r, fig.height=8}
bronx_data <- airbnb_clean |>
  filter(neighbourhood_group == "Bronx") |>
  group_by(neighbourhood) |>
  filter(n() >= 10) |>
  ungroup()

if (nrow(bronx_data) > 0) {
  ggplot(bronx_data, aes(x = reorder(neighbourhood, price, FUN = median), y = price)) +
    geom_boxplot(fill = "#BC4749", alpha = 0.8, outlier.alpha = 0.2) +
    scale_y_log10(labels = dollar_format()) +
    coord_flip() +
    labs(
      title = "Bronx Neighborhood Pricing Overview",
      subtitle = "Entry-level market opportunities and neighborhood dynamics",
      x = NULL,
      y = "Nightly Rate (USD, log scale)",
      caption = "Figure 3.4"
    ) +
    theme(axis.text.y = element_text(size = 7))
}
```

## Staten Island — Niche Market Analysis

```{r, fig.height=8}
staten_data <- airbnb_clean |>
  filter(neighbourhood_group == "Staten Island") |>
  group_by(neighbourhood) |>
  filter(n() >= 10) |>
  ungroup()

if (nrow(staten_data) > 0) {
  ggplot(staten_data, aes(x = reorder(neighbourhood, price, FUN = median), y = price)) +
    geom_boxplot(fill = "#FF385C", alpha = 0.8, outlier.alpha = 0.2) +
    scale_y_log10(labels = dollar_format()) +
    coord_flip() +
    labs(
      title = "Staten Island Neighborhood Market Profile",
      subtitle = "Specialized market with limited but distinct offerings",
      x = NULL,
      y = "Nightly Rate (USD, log scale)",
      caption = "Figure 3.5"
    ) +
    theme(axis.text.y = element_text(size = 7))
}
```

---

# Predictive Modeling Framework

<div class="result-card" style="background: linear-gradient(135deg, #f8fafc 0%, #f1f5f9 100%); border-top: 4px solid #2563eb;">

## Model Objective {-}

Develop and validate machine learning models to predict Airbnb nightly rates based on **location**, **property characteristics**, and **demand indicators**. Two complementary approaches are evaluated:

- **Linear Regression:** Interpretable coefficient-based model for understanding factor contributions
- **Regression Tree:** Non-parametric model capturing non-linear relationships and interactions

**Business Value:** Enable data-driven pricing strategies and investment targeting through accurate price forecasting.

</div>

## Data Preparation Protocol

```{r}
# Feature engineering and train-test split
airbnb_model <- airbnb_clean |>
  mutate(
    log_price = log1p(price),
    room_type = as.factor(room_type),
    neighbourhood_group = as.factor(neighbourhood_group),
    number_of_reviews = as.numeric(number_of_reviews)
  )

set.seed(42)
n <- nrow(airbnb_model)
idx <- sample.int(n, size = floor(0.7*n))
train <- airbnb_model[idx, ]
test  <- airbnb_model[-idx, ]
```

<div class="kpi-row">

<div class="metric">
  <span class="metric-value">`r format(nrow(train), big.mark = ",")`</span>
  <span class="metric-label">Training Observations</span>
</div>

<div class="metric">
  <span class="metric-value">`r format(nrow(test), big.mark = ",")`</span>
  <span class="metric-label">Test Observations</span>
</div>

<div class="metric">
  <span class="metric-value">70/30</span>
  <span class="metric-label">Split Ratio</span>
</div>

</div>

## Model 1: Multiple Linear Regression

```{r}
# Train linear model
lm_fit <- lm(log_price ~ neighbourhood_group + room_type + number_of_reviews, data = train)
lm_sum <- summary(lm_fit)
```

<div class="success-box">

**Model Performance:**

- **R² Score:** `r round(lm_sum$r.squared, 4)` (`r round(lm_sum$r.squared * 100, 2)`% variance explained)
- **Adjusted R²:** `r round(lm_sum$adj.r.squared, 4)`
- **Residual Std Error:** `r round(lm_sum$sigma, 4)`
- **F-Statistic:** `r round(lm_sum$fstatistic[1], 2)` (p < 0.001)

The model demonstrates statistically significant explanatory power, with location and property type as primary determinants.

</div>

### Coefficient Interpretation

```{r, fig.height=2, fig.width=3}
# Extract and transform coefficients
coefs <- coefficients(lm_fit)
effects_price <- expm1(coefs)

percentOfInfluence <- c(
  effects_price["neighbourhood_groupBrooklyn"] * 100,
  effects_price["neighbourhood_groupManhattan"] * 100,
  effects_price["neighbourhood_groupQueens"] * 100,
  effects_price["neighbourhood_groupStaten Island"] * 100,
  effects_price["room_typePrivate room"] * 100,
  effects_price["room_typeShared room"] * 100
)

coef_data <- tibble(
  Variable = c(
    "Brooklyn",
    "Manhattan",
    "Queens",
    "Staten Island",
    "Private Room",
    "Shared Room"
  ),
  Effect = percentOfInfluence
) |>
  arrange(desc(Effect))

# Calculate range for axis limits
max_abs_effect <- max(abs(coef_data$Effect))
x_range <- c(-max_abs_effect * 1.25, max_abs_effect * 1.25)

# Create interactive plotly chart
plot_ly(coef_data, 
        x = ~Effect, 
        y = ~reorder(Variable, Effect),
        type = 'bar',
        orientation = 'h',
        marker = list(color = ~ifelse(Effect > 0, '#667eea', '#f093fb')),
        text = ~paste0(ifelse(Effect > 0, "+", ""), round(Effect, 1), "%"),
        textposition = 'outside',
        hovertemplate = '<b>%{y}</b><br>Effect: %{x:.2f}%<extra></extra>') |>
  layout(
    title = list(
      text = "Price Impact by Factor<br><sub>Percentage change vs baseline (Bronx, Entire home/apt)</sub>",
      font = list(size = 14, color = '#1a2332')
    ),
    xaxis = list(
      title = "Effect on Price (%)",
      titlefont = list(size = 11, color = '#1a2332'),
      gridcolor = '#e5e5e5',
      range = x_range
    ),
    yaxis = list(
      title = "",
      titlefont = list(size = 11, color = '#1a2332')
    ),
    plot_bgcolor = 'rgba(0,0,0,0)',
    paper_bgcolor = 'rgba(0,0,0,0)',
    margin = list(l = 120, r = 100, t = 80, b = 50),
    autosize = TRUE
  )
```

## Model 2: Pruned Regression Tree

```{r}
# Train and optimize decision tree
tree_fit <- rpart(
  log_price ~ neighbourhood_group + room_type + number_of_reviews,
  data = train,
  method = "anova",
  control = rpart.control(cp = 0.01, xval = 10)
)

best_cp <- tree_fit$cptable[which.min(tree_fit$cptable[, "xerror"]), "CP"]
tree_fit_pruned <- prune(tree_fit, cp = best_cp)
```

### Variable Importance Ranking

```{r}
# Extract feature importance
importance_data <- tibble(
  Variable = names(tree_fit_pruned$variable.importance),
  Importance = tree_fit_pruned$variable.importance
) |>
  mutate(
    `Relative Importance (%)` = round(Importance / sum(Importance) * 100, 1),
    Rank = row_number()
  ) |>
  select(Rank, Variable, `Relative Importance (%)`) |>
  arrange(Rank)

kable(
  importance_data,
  caption = "**Table 4.2** — Decision Tree Variable Importance Hierarchy",
  align = 'clc'
)
```

<div class="success-box">

**Key Finding:**

**Room type** dominates the tree's decision structure, accounting for **~75%** of total variable importance. This finding validates the linear model's coefficients and confirms that property type is the single most critical pricing determinant, surpassing geographical factors.

</div>

## Model Validation & Performance

```{r}
# Generate predictions
pred_log_lm_test <- predict(lm_fit, newdata = test)
pred_price_lm_test <- expm1(pred_log_lm_test)

pred_log_tree_test <- predict(tree_fit_pruned, newdata = test)
pred_price_tree_test <- expm1(pred_log_tree_test)

# Calculate RMSE
rmse <- function(x, y) sqrt(mean((x - y)^2, na.rm = TRUE))

rmse_lm <- rmse(pred_price_lm_test, test$price)    # Dollars Scale
rmse_tree <- rmse(pred_price_tree_test, test$price)

rmse_lm_log <- rmse(pred_log_lm_test, test$log_price)   # Log Scale
rmse_tree_log <- rmse(pred_log_tree_test, test$log_price)

# Create comparison table
model_comparison <- tibble(
  Model = c("Linear Regression", "Regression Tree"),
  `RMSE (Log Scale)` = round(c(rmse_lm_log, rmse_tree_log), 2),
  `RMSE ($)` = round(c(rmse_lm, rmse_tree), 2),
  `RMSE (%)` = paste0(round(c(rmse_lm / mean(test$price) * 100, 
                              rmse_tree / mean(test$price) * 100), 1), "%")
)

kable(
  model_comparison,
  caption = "**Table 4.3** — Model Performance Benchmarking on Test Set",
  align = 'lccc'
)
```

<div class="kpi-row">

<div class="metric">
  <span class="metric-value">$`r round(rmse_lm, 0)`</span>
  <span class="metric-label">Linear Regression RMSE</span>
</div>

<div class="metric">
  <span class="metric-value">$`r round(rmse_tree, 0)`</span>
  <span class="metric-label">Regression Tree RMSE</span>
</div>

<div class="metric">
  <span class="metric-value">`r round((1 - rmse_lm / mean(test$price)) * 100, 1)`%</span>
  <span class="metric-label">Accuracy Rate</span>
</div>

</div>

### Performance Assessment

<div class="highlight-box">

**Model Evaluation Summary:**

Both models achieve comparable performance with RMSE values of approximately **$`r round(rmse_lm, 0)`**. While this represents meaningful predictive capability, the error magnitude remains significant relative to median pricing ($`r round(median(test$price), 0)`).

**Limitations Identified:**

- Current feature set captures ~`r round(lm_sum$r.squared * 100, 0)`% of price variance
- Additional variables (amenities, ratings, seasonality) required for enhanced accuracy
- Model performs better on mid-tier properties than luxury or budget segments

**Recommended Next Steps:**

1. Enrich feature space with amenity data and host quality metrics
2. Implement ensemble methods (Random Forest, Gradient Boosting)
3. Develop segment-specific models for different price ranges
4. Incorporate temporal features for seasonal pricing optimization

</div>

---

# Conclusions & Strategic Recommendations

## Executive Summary of Findings

1. **Market Structure:** Manhattan and Brooklyn dominate with 80%+ market share; significant pricing premiums observed
2. **Property Dynamics:** Entire homes command 2-3x premiums over private rooms; shared rooms represent minimal market opportunity
3. **Pricing Drivers:** Location (borough) and property type explain 60%+ of price variation
4. **Demand Signals:** Inverse correlation between availability and review volume indicates supply-demand imbalance in premium segments

## Strategic Implications

<div class="success-box">

**Investment Targeting:** Focus acquisition strategies on Manhattan (premium positioning) and Brooklyn (growth/value opportunities)

**Pricing Strategy:** Room type should be primary pricing lever; neighborhood-level optimization secondary but meaningful

**Portfolio Optimization:** Prioritize entire home/apartment inventory; limited-availability positioning enhances perceived value

**Future Analytics:** Integrate additional data sources (seasonal trends, competitive set analysis, amenity indexes) to refine forecasting accuracy

</div>

---

<div style="text-align: center; padding: 60px 40px; margin-top: 60px; border-top: 3px solid #e5e5e5; background: #fafafa;">

**New York City Airbnb Market Analysis**

Made by Antoine, Charles, Mokhmad

</div>